Empirical Bayes estimation of the normal inverse Gaussian prior in simultaneous-equation models

### 2019-04-24 docs/manuscript1.Rnw
first two simulations ran, intercept added, split functions.R

% settings
\graphicspath{{../figs/}}
\pgfplotsset{compat=1.16}
\setlength{\evensidemargin}{.5cm} \setlength{\oddsidemargin}{.5cm}
\setlength{\textwidth}{15cm}
\title{Empirical Bayes estimation of simultaneous-equation models}
\date{\today}
\author{Magnus M. M\"unch$^{1,2}$\footnote{Correspondence to: 
\href{mailto:m.munch@vumc.nl}{m.munch@vumc.nl}}, Mark A. van de Wiel$^{1,3}$, 
Sylvia Richardson$^{3}$, \\ and Gwena{\"e}l G. R. Leday$^{3}$}

@@

	\noindent\textbf{Keywords}: Empirical Bayes; Simultaneous-equations model; 
	Variational Bayes; Multivariate data

@@

	- Regression information

@@

	- our solution: adaptive borrowing of information through empirical Bayes, 
	guided by external regression information.

@@

	\subsection{Bayesian prior model}
	We capture the uncertainty in the parameters through a Bayesian prior model. 
	We assign a flat prior to the intercept terms $\beta_{0d}$ and integrate 
	them out. The remaining parameters are endowed with the following priors:
	\begin{subequations}\label{eq:priormodel}
	  \begin{align}
      y_{id} | \bm{\beta}_d, \sigma_d^2 & \sim \mathcal{N} 
      (\x_i \tr \bm{\beta}_d, \sigma_d^2), \\
      \beta_{jd} | \gamma_d^2, \sigma_d^2 & \sim 
      \mathcal{N} (0, \sigma_d^2 \gamma_d^2), \label{eq:betaprior} \\
      \gamma_d^2 & \sim \mathcal{IG}(\theta_d, \lambda_d), \\
      \sigma_d^2 & \sim 1/\sigma_d^3,
    \end{align}
  \end{subequations}
  where $\mathcal{IG}(\theta, \lambda)$ denotes the inverse Gaussian 
  distribution with mean $\theta$ and shape $\lambda$.
  \cite{moran_variance_2018} argue that in regression models, $\bm{\beta}_d$ 
  and $\sigma_d^2$ should be independent \textit{a priori}, i.e.
  $\beta_{jd} | \gamma_d^2, \sigma_d^2 \sim \beta_{jd} | 
  \gamma_d^2$. However, in our case
  that is debatable. The $D$ regressions need not be on the same scale, so in
  order to jointly consider the $\bm{\beta}_d$ some 
  form of calibration is required. A straightforward way 
  of doing this is to include the scale the $\bm{\beta}_d$ prior with 
  $\sigma_d^2$ as in (\ref{eq:betaprior}).

  The normal inverse Gaussian (NIG) prior model for $\beta_{jd}$ and 
  $\gamma^2_d$ in (\ref{eq:priormodel}) was introduced in 
  \cite{barndorff-nielsen_hyperbolic_1978} and is, since 
  \cite{barndorff-nielsen_normal_1997}, routinely applied in mathematical 
  finance \cite[]{kalemanova_normal_2007}. In Figure 
  \ref{fig:dens_igaussian_marginalbeta}, the NIG 
  prior distribution is depicted together with two common prior choices, ridge 
  and lasso. From this Figure we see that, depending on the choice
  of hyperparameters, the NIG prior can vary from heavier tails than the lasso, 
  to more Gaussian tails like the ridge.
<<dens_igaussian_marginalbeta, fig.cap="Marginal $\\beta_j$ prior for several choices of $\\gamma^2_d$ hyperpior, scaled to $\\V(\\beta_j)=1$.", out.width="60%", fig.asp=1>>=
library(GeneralizedHyperbolic)
# all scaled such that variance is one
sigma <- 1
theta <- 1/sigma^2
lambda <- c(0.1, 2)
lambda1 <- sqrt(2)
col <- c(1:4)
lty <- c(1:4)
labels <- as.expression(c(bquote("IG, "~lambda==.(lambda[1])~", "~
                                   theta==.(theta)),
                          bquote("IG, "~lambda==.(lambda[2])~", "~
                                   theta==.(theta)),
                          "Point mass (ridge)", "Exponential (lasso)"))
dprior <- function(x, lambda, theta, sigma) {
  dnig(x, 0, sigma*sqrt(lambda), sqrt(lambda/(theta*sigma)), 0)
}
dlasso <- function(x, lambda1) {
  0.5*lambda1*exp(-lambda1*abs(x))
}

curve(dprior(x, lambda[1], theta, sigma), -3, 3, ylab=expression(p(beta[j])), 
      xlab=expression(beta[j]), n=1000, 
      col=col[1], lty=lty[1])
curve(dprior(x, lambda[2], theta, sigma), add=TRUE, n=1000, 
      col=col[2], lty=lty[2])
curve(dnorm(x, 0, 1), add=TRUE, n=1000, col=col[3], lty=lty[3])
curve(dlasso(x, lambda1), add=TRUE, n=1000, col=col[4], lty=lty[4])
legend("topright", legend=labels, col=col, lty=lty, 
       title="Hyperprior")
@

  Our argumentation to model the $\gamma^2_d$ by an inverse Gaussian 
  distribution, as has been suggested in \cite{fabrizi_specification_2016} and 
  \cite{caron_sparse_2008}, is two-fold: (i) the NIG model is more flexible than
  the standard ridge and (ii) allows to model the mean $\theta_d$ as a 
  function of the co-data $\mathbf{c}_d$, as explained in Section 
  \ref{sec:empiricalbayes}.

  A few remarks on the choice of error variance prior are justified here: 
  many authors endow error variance components with vague gamma priors. 
  \cite{gelman_prior_2006}, 
  among others, advises against this practice. The degree of `vagueness' has a 
  large influence on the posterior, while degree of `vagueness' is a difficult 
  parameter to set. This influence is especially pronounced if the likelihood is
  relatively flat, as may be reasonably expected in the large $p$, small $n$ 
  setting we are in. We therefore model the error variance with Jeffrey's prior 
  \cite[]{jeffreys_invariant_1946}. In the derivation of our Jeffrey's prior
  for the error variance, we jointly consider an unknown mean and variance.

@@

	\section{Estimation}
	\subsection{Variational Bayes}\label{sec:variationalbayes}
	The posterior corresponding to the model described in (\ref{eq:linearmodel})
	and (\ref{eq:priormodel}) is not available in closed form. We therefore 
	approximate the posterior by variational Bayes, where we force the posterior 
	density $Q_d$ to factorise as: $p(\bbeta_d, \sigma_d^2, \gamma_d^2) \approx 
	Q_d(\cdot) = q(\bbeta_d) \cdot q(\sigma_d^2) \cdot q(\gamma_d^2)$.
	Under such a factorisation, the marginal variational posteriors that 
	minimise the Kullback-Leibler divergence of the true posterior to the 
	variational Bayes approximation \cite[]{neal_view_1998}, are given by:
	\begin{align*}
    q(\bm{\beta}_d) & \overset{D}{=} \mathcal{N}_p 
    (\bm{\mu}_d, \bm{\Sigma}_d), \\
    q(\gamma^2_d) & \overset{D}{=} \mathcal{GIG}
    \(-\frac{p+1}{2}, \lambda_d \theta_d^{-2}, \delta_d\), \\
    q(\sigma^2_d) & \overset{D}{=} \Gamma^{-1} \(\frac{n + p + 1}{2}, \zeta_d\).
  \end{align*}
	The variational parameters $\bm{\mu}_d$, $\bm{\Sigma}_d$, $\delta_d$, and
	$\zeta_d$ contain cyclic dependencies and are iteratively 
	updated by:
  \begin{align*}
    \bm{\Sigma}_d^{(h+1)} & = (a_d^{(h)})^{-1} 
    (\X \tr \X + b_d^{(h)} \I)^{-1}, \\
    \bm{\mu}_d^{(h+1)} & = (\X \tr \X + b_d^{(h)} \I)^{-1} \X \tr \y_d, \\
    \delta_d^{(h+1)} & = \bmu_d \tr \bmu_d + \lambda_d, \\ 
    \zeta_d^{(h+1)} & = \frac{1}{2} \bigg[\mathbf{y}_d \tr \mathbf{y}_d -
    2 \mathbf{y}_d \tr \X \bm{\mu}_d^{(h+1)} + 
    \trace ( \X \tr \X \bm{\Sigma}_d^{(h+1)}) + 
    (\bm{\mu}_d^{(h+1)}) \tr \X \tr \X \bm{\mu}_d^{(h+1)} \\
    & \,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\, + 
    b_d^{(h+1)} \trace ( \bm{\Sigma}_d^{(h+1)}) + 
    b_d^{(h+1)} (\bm{\mu}_d^{(h+1)}) \tr \bm{\mu}_d^{(h+1)}\bigg],
  \end{align*}
  until convergence. Here, we set
  \begin{align*}
    a_d^{(h)} & =\E_{Q^{(h)}}(\sigma_d^{-2})=(n + p + 1)/(2 \zeta_d^{(h)}), \\
    b_d^{(h)} & = \E_{Q^{(h)}}(\gamma_d^{-2}) = \sqrt{\frac{\lambda_d}
    {\theta_d^2 \delta_d^{(h)}}} \frac{K_{\frac{p - 1}{2}} 
    \( \sqrt{\lambda_d \delta_d^{(h)}}/\theta_d \)}
    {K_{\frac{p+1}{2}} \( \sqrt{\lambda_d \delta_d^{(h)}}/\theta_d \)} + 
    \frac{p+1}{\delta_d^{(h)}}.
  \end{align*}

@@

	To make use of the outcome-specific codata, we parametrise 
	$\theta_d = (\mathbf{c}_d \tr \bm{\alpha})^{-1}$ and estimate $\blambda$ and
	$\bm{\alpha}$. This allows us to include continuous co-data on the outcomes
	in an objective, data-driven manner. The estimating equations are:
  \begin{align*}
    \bm{\alpha} & = \[ \mathbf{C} \tr \diag 
    (\lambda_d e_d^{(l)}) \mathbf{C} \]^{-1} \mathbf{C} \tr \diag 
    (\lambda_d) \mathbf{1}_{D \times 1}, \\
    \lambda_d & = \[ b_d^{(l)} + e_d^{(l)} (\mathbf{c}_d \tr \bm{\alpha})^2 - 
    2\mathbf{c}_d \tr \bm{\alpha} \]^{-1},
  \end{align*}
  where $e_d^{(l)} = \E_{Q^{(l)}}(\gamma_d^2) = [b_d^{(l)} - 
  (p + 1)/\delta_d^{(l)})] \cdot \delta_d^{(l)} 
  (\theta_d^{(h)})^2/\lambda_d^{(l)}$. Solving the equations is done by 
  iteratively reweighted least squares of responses $(e_d^{(l)})^{-1}$ on 
  predictors $\mathbf{c}_d$ with weights $\lambda_d e_d^{(l)}$.

  \section{Comparison of NIG model to Normal inverse Gamma}
  - check model with Stan.

  \section{Comparison of conjugate to non-conjugate}

  \section{Variational Bayes versus MCMC}

  \section{Inverse Gaussian Empirical Bayes versus inverse Gamma}

  \section{Simulations for VB}
	\subsection{Setup}

  We simulate from a correctly specified model given in \ref{eq:priormodel}. We assume random $y_{id}$, $\x_i$, and $\bm{\beta}_d$, while we fix $\mathbf{C}$, $\gamma_d^2$, $\sigma_d^2$ and $\bm{\alpha}$. We estimate the $\bm{\mu}_d$, $\bm{\Sigma}_d$, $\delta_d$, $\bm{\alpha}$ and $\lambda$, while we fix the $\sigma_d^2$ to their true values.

  We set $n=100$, $p=200$, and $D=10$. We consider 5 evenly sized classes of drugs, such that $\mathbf{c}_{kd} = \mathbbm{1}(k = \text{class}_d)$ and $\begin{bmatrix} \mathbf{c}_1 & \cdots & \mathbf{c}_D \end{bmatrix} \tr$. From these the $\gamma_d^2 = (\mathbf{c}_d \tr \bm{\alpha})^{-1}$ are created. Next, we simulate $\bm{\beta}_{jd} \sim \mathcal{N} (0, \sigma^2_d \gamma_d^2)$. We fix the mean signal-to noise ratio $\overline{\text{SNR}}_d$, such that the predictor data variance is $s^2 = \V(\mathbf{x}_i) = \overline{\text{SNR}}_d / (\overline{\gamma^2_d} p)$ (NOT CORRECT). The last step consists of simulating the data by $\mathbf{x}_{ij} \sim \mathcal{N}(0,s^2)$ and $y_{id} = \mathcal{N} (\mathbf{x}_i \tr \bm{\beta}_d, \sigma_d^2)$. The remaining parameters are set as follows: $\bm{\alpha} = \begin{bmatrix} 1 & \dots & 5 \end{bmatrix} \tr$, $\sigma_d^2 = 1$ for all $d$, and $\overline{\text{SNR}}_d = 10$. To account for the random data and parameter generation, we repeat the simulation 100 times. We compare the results to the inverse Gamma and non-conjugate model, and present the results in Figures \ref{fig:boxplot_igaussian_res1_prior_mean}-\ref{fig:boxplot_igaussian_res2_prior_mean}.

@@

@@

  \section{Discussion}

>>>>

% settings
\graphicspath{{../figs/}}
\pgfplotsset{compat=1.16}
\setlength{\evensidemargin}{.5cm} \setlength{\oddsidemargin}{.5cm}
\setlength{\textwidth}{15cm}
\title{Empirical Bayes estimation of the normal inverse Gaussian prior in 
simultaneous-equation models}
\date{\today}
\author{Magnus M. M\"unch$^{1,2}$\footnote{Correspondence to: 
\href{mailto:m.munch@vumc.nl}{m.munch@vumc.nl}}, Mark A. van de Wiel$^{1,3}$, 
Sylvia Richardson$^{3}$, \\ and Gwena{\"e}l G. R. Leday$^{3}$}

@@

	\noindent\textbf{Keywords}: ...

@@

	- Co-data introduction

@@

	- our solution: adaptive borrowing of information through empirical Bayes, 
	guided by co-data.

@@

	\subsection{Bayesian prior model}
	We capture the uncertainty in the parameters through a Bayesian prior model. 
	We assign a flat prior to the intercept terms $\beta_{0d}$ and integrate 
	them out. The remaining parameters are endowed with the following priors:
	\begin{subequations}\label{eq:priormodel}
	  \begin{align}
      y_{id} | \bm{\beta}_d, \sigma_d^2 & \sim \mathcal{N} 
      (\x_i \tr \bm{\beta}_d, \sigma_d^2), \\
      \beta_{jd} | \gamma_d^2, \sigma_d^2 & \sim 
      \mathcal{N} (0, \sigma_d^2 \gamma_d^2), \label{eq:betaprior} \\
      \gamma_d^2 & \sim \mathcal{IG}(\theta_d, \lambda_d), \\
      \sigma_d^2 & \sim 1/\sigma_d^3,
    \end{align}
  \end{subequations}
  where $\mathcal{IG}(\theta, \lambda)$ denotes the inverse Gaussian 
  distribution with mean $\theta$ and shape $\lambda$.
  \cite{moran_variance_2018} argue that in regression models, $\bm{\beta}_d$ 
  and $\sigma_d^2$ should be independent \textit{a priori}, i.e.
  $\beta_{jd} | \gamma_d^2, \sigma_d^2 \sim \beta_{jd} | 
  \gamma_d^2$. However, in our case
  that is debatable. The $D$ regressions need not be on the same scale, so in
  order to jointly consider the $\bm{\beta}_d$ some 
  form of calibration is required. A straightforward way 
  of doing this is to scale the $\bm{\beta}_d$ prior with 
  $\sigma_d^2$ as in (\ref{eq:betaprior}), resulting in a 
  scale-free prior with respect to the response.

  The normal inverse Gaussian (NIG) prior model for $\beta_{jd}$ and 
  $\gamma^2_d$ in (\ref{eq:priormodel}) was introduced in 
  \cite{barndorff-nielsen_hyperbolic_1978} and is, since 
  \cite{barndorff-nielsen_normal_1997}, routinely applied in mathematical 
  finance (see, e.g., \cite{kalemanova_normal_2007}). In Figure 
  \ref{fig:dens_igaussian_marginalbeta}, the NIG 
  prior distribution marginalised over $\gamma_d^2$ is depicted together with 
  three other common prior choices,
  Student's $t$ (see Section \ref{sec:inversegammamodel} for more details on 
  this prior), ridge and lasso, all scaled to variance one. From this Figure 
  we see that, depending on the choice
  of hyperparameters, the NIG prior can vary from heavier tails than the lasso, 
  to more Gaussian tails like the ridge, but behaves similarly to the 
  Student's $t$ prior. Our argumentation to model the 
  $\gamma^2_d$ by an inverse Gaussian 
  distribution, as has been suggested in \cite{fabrizi_specification_2016} and 
  \cite{caron_sparse_2008}, is two-fold: (i) the NIG model is more flexible than
  the standard ridge and lasso priors (as seen from Figure 
  \ref{fig:dens_igaussian_marginalbeta}), and (ii) the NIG prior allows to model
  the mean $\theta_d$ as a 
  function of the co-data $\mathbf{c}_d$ more conveniently than the Student's 
  $t$ prior, as explained in Section 
  \ref{sec:empiricalbayes}. 
<<dens_igaussian_marginalbeta, fig.cap="Several choices of marginal $\\beta_j$ prior, scaled to $\\V(\\beta_j)=1$.", out.width="80%", fig.asp=2/3>>=
library(GeneralizedHyperbolic)
library(metRology)
# all scaled such that variance is one
sigma <- 1
theta <- 1/sigma^2
lambda <- c(0.1, 2)
lambda1 <- sqrt(2)
eta <- lambda/theta + 2
col <- c(1:6)
lty <- c(1:6)
labels <- as.expression(c(bquote("NIG, "~lambda==.(lambda[1])~", "~
                                   theta==.(theta)),
                          bquote("NIG, "~lambda==.(lambda[2])~", "~
                                   theta==.(theta)),
                          bquote("Student's t, "~lambda==.(lambda[1])~", "~
                                   eta==.(eta[1])),
                          bquote("Student's t, "~lambda==.(lambda[2])~", "~
                                   eta==.(eta[2])),
                          "ridge", "lasso"))
dprior <- function(x, lambda, theta, sigma) {
  dnig(x, 0, sigma*sqrt(lambda), sqrt(lambda/(theta*sigma)), 0)
}
dlasso <- function(x, lambda1) {
  0.5*lambda1*exp(-lambda1*abs(x))
}

ylim <- c(0, dt.scaled(0, eta[1]/2, 0, sqrt(lambda[1]*sigma^2/eta[1])))
curve(dprior(x, lambda[1], theta, sigma), -3, 3, ylab=expression(p(beta[j])), 
      xlab=expression(beta[j]), n=1000, ylim=ylim,
      col=col[1], lty=lty[1])
curve(dprior(x, lambda[2], theta, sigma), add=TRUE, n=1000, 
      col=col[2], lty=lty[2])
curve(dnorm(x, 0, 1), add=TRUE, n=1000, col=col[3], lty=lty[3])
curve(dlasso(x, lambda1), add=TRUE, n=1000, col=col[4], lty=lty[4])
curve(dt.scaled(x, eta[1]/2, 0, sqrt(lambda[1]*sigma^2/eta[1])), add=TRUE, 
      n=1000, col=col[5], lty=lty[5])
curve(dt.scaled(x, eta[2]/2, 0, sqrt(lambda[2]*sigma^2/eta[2])), add=TRUE, 
      n=1000, col=col[6], lty=lty[6])
legend("topright", legend=labels, col=col, lty=lty, 
       title="Prior", seg.len=1)
@

  A few remarks on the choice of error variance prior are justified here: 
  many authors endow error variance components with vague gamma priors. 
  \cite{gelman_prior_2006}, 
  among others, advises against this practice. The degree of `vagueness' has a 
  large influence on the posterior, while degree of `vagueness' is a difficult 
  parameter to set. This influence is especially pronounced if the likelihood is
  relatively flat, as may be reasonably expected in the large $p$, small $n$ 
  setting we are in. We therefore model the error variance with Jeffrey's 
  objective prior 
  \cite[]{jeffreys_invariant_1946}. In the derivation of our Jeffrey's prior
  for the error variance, we jointly consider an unknown mean and variance.

@@

	\section{Estimation}
	\subsection{Variational Bayes}\label{sec:variationalbayes}
	The posterior corresponding to the model described in (\ref{eq:linearmodel})
	and (\ref{eq:priormodel}) is not available in closed form. We therefore 
	approximate the posterior by variational Bayes, where we force the approximate
	posterior 
	density to factorise as: $p(\bbeta_d, \sigma_d^2, \gamma_d^2) \approx 
	Q_d(\cdot) = q(\bbeta_d) \cdot q(\sigma_d^2) \cdot q(\gamma_d^2)$. For
	notational convenience, we slightly abuse notation and let $q(\cdot)$ denote
	different densities for different inputs.
	Under such a factorisation, the marginal variational posteriors that 
	minimise the Kullback-Leibler divergence of the true posterior to the 
	variational Bayes approximation \cite[]{neal_view_1998} are given by:
	\begin{align*}
    q(\bm{\beta}_d) & \overset{D}{=} \mathcal{N}_p 
    (\bm{\mu}_d, \bm{\Sigma}_d), \\
    q(\gamma^2_d) & \overset{D}{=} \mathcal{GIG}
    \(-\frac{p+1}{2}, \lambda_d \theta_d^{-2}, \delta_d\), \\
    q(\sigma^2_d) & \overset{D}{=} \Gamma^{-1} \(\frac{n + p + 1}{2}, \zeta_d\).
  \end{align*}
	See SM Section \ref{sm-sec:vbderivations} for the derivations. The variational
	parameters $\bm{\mu}_d$, $\bm{\Sigma}_d$, $\delta_d$, and
	$\zeta_d$ contain cyclic dependencies and are iteratively 
	updated by:
  \begin{align*}
    \bm{\Sigma}_d^{(h+1)} & = (a_d^{(h)})^{-1} 
    (\X \tr \X + b_d^{(h)} \I)^{-1}, \\
    \bm{\mu}_d^{(h+1)} & = (\X \tr \X + b_d^{(h)} \I)^{-1} \X \tr \y_d, \\
    \delta_d^{(h+1)} & = a_d^{(h)} \[(\bmu^{(h+1)}_d) \tr \bmu^{(h+1)}_d +
    \trace (\bSigma_d^{(h+1)})\] + \lambda_d, \\ 
    \zeta_d^{(h+1)} & = \frac{1}{2} \bigg[\mathbf{y}_d \tr \mathbf{y}_d -
    2 \mathbf{y}_d \tr \X \bm{\mu}_d^{(h+1)} + 
    \trace ( \X \tr \X \bm{\Sigma}_d^{(h+1)}) + 
    (\bm{\mu}_d^{(h+1)}) \tr \X \tr \X \bm{\mu}_d^{(h+1)} \\
    & \,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\, + 
    b_d^{(h+1)} \trace ( \bm{\Sigma}_d^{(h+1)}) + 
    b_d^{(h+1)} (\bm{\mu}_d^{(h+1)}) \tr \bm{\mu}_d^{(h+1)}\bigg],
  \end{align*}
  until convergence. Here, we set
  \begin{align}
    a_d^{(h)} & =\E_{Q^{(h)}}(\sigma_d^{-2})=(n + p + 1)/(2 \zeta_d^{(h)}), 
    \nonumber\\
    b_d^{(h)} & = \E_{Q^{(h)}}(\gamma_d^{-2}) = \sqrt{\frac{\lambda_d}
    {\theta_d^2 \delta_d^{(h)}}} \frac{K_{\frac{p - 1}{2}} 
    \( \sqrt{\lambda_d \delta_d^{(h)}}/\theta_d \)}
    {K_{\frac{p+1}{2}} \( \sqrt{\lambda_d \delta_d^{(h)}}/\theta_d \)} + 
    \frac{p+1}{\delta_d^{(h)}}, \label{eq:ratiomodifiedbessel}
  \end{align}
  where $K_{\nu}(x)$ denotes the modified Bessel function of the second kind.
  A method for fast and numerically stable calculation of ratios of such 
  functions as in (\ref{eq:ratiomodifiedbessel}) is given in SM Section 
  \ref{sm-sec:ratiosmodifiedbessels}.

@@

	To make use of the outcome-specific codata, we parametrise 
	$\theta_d = (\mathbf{c}_d \tr \bm{\alpha})^{-1}$ and estimate $\blambda$ and
	$\bm{\alpha}$. This allows us to include continuous co-data on the outcomes
	in an objective, data-driven manner. Additionally, it reduces the number of
	hyperparameters even further. The estimating equations are:
  \begin{align*}
    \bm{\alpha} & = \[ \mathbf{C} \tr \diag 
    (\lambda_d e_d^{(l)}) \mathbf{C} \]^{-1} \mathbf{C} \tr \diag 
    (\lambda_d) \mathbf{1}_{D \times 1}, \\
    \lambda_d & = \[ b_d^{(l)} + e_d^{(l)} (\mathbf{c}_d \tr \bm{\alpha})^2 - 
    2\mathbf{c}_d \tr \bm{\alpha} \]^{-1},
  \end{align*}
  where $e_d^{(l)} = \E_{Q^{(l)}}(\gamma_d^2) = [b_d^{(l)} - 
  (p + 1)/\delta_d^{(l)})] \cdot \delta_d^{(l)} 
  (\theta_d^{(h)})^2/\lambda_d^{(l)}$. Solving the equations is done by 
  iteratively reweighted least squares of responses $(e_d^{(l)})^{-1}$ on 
  predictors $\mathbf{c}_d$ with weights $\lambda_d e_d^{(l)}$.

  Variaional Bayes approximations are known to underestimate posterior 
  variances \cite[]{rue_approximate_2009,consonni_mean-field_2007,
  bishop_pattern_2006,wang_inadequacy_2005}. If posterior variances are needed, 
  we suggest generating samples from the posterior after the empirical Bayes 
  iterations have converged. A Gibbs sampler to do so is described in 
  the SM Section \ref{sm-sec:gibbssampler}. Alternatively, the \texttt{R} 
  package \texttt{rstan} \cite[]{guo_rstan:_2018} allows the user to sample
  from a generic posterior. An implementation of model
  (\ref{eq:priormodel}) in \texttt{stan} code may be found on 
  \url{https://github.com/magnusmunch/cambridge}.

  \section{Simulations}
  \subsection{Comparison of NIG to normal inverse Gamma model}
  \label{sec:inversegammamodel}
  A common default prior for latent variance components is the inverse Gamma
  prior:
  $$
  \gamma_d^2 \sim \Gamma^{-1}(\eta_d/2,\lambda_d/2),
  $$
  with shape $\eta_d/2$ and scale $\lambda_d/2$, both rescaled to match our 
  inverse Gaussian prior formulation. The resulting marginal prior for the model
  parameters
  (conditional on the error variance $\sigma_d^2$) is then a scaled Student's 
  $t$-distribution:
  $$
  \beta_j | \sigma_d^2 \sim \text{Student's } 
  t\(\eta_d,\sigma_d \sqrt{\lambda_d/\eta_d}\),
  $$
  with degrees of freedom $\eta_d$ and scale $\sigma_d \sqrt{\lambda_d/\eta_d}$.
  See SM Section \ref{sm-sec:inversegamma} for more details on this prior. 
  We compare the 
  posteriors of the proposed NIG prior to the scaled Student's $t$-distribution
  prior in a simulated setting. We generate draws from these posteriors using 
  an MCMC sampler implemented in \texttt{rstan}. The aim of this simulation 
  study is 
  to investigate the performance of the NIG prior model, so we set $D=1$ and 
  drop the $d$ indices in the following.

  \subsection{Variational Bayes approximation}
  To investigate the accuracy of our variational Bayes approximation we compare 
  the resulting posterior to an MCMC sample from the posterior obtained with 
  \texttt{rstan}. Since the MCMC
  sample represents the exact posterior (up to Monte Carlo error), we use it as 
  a proxy
  for the true posterior. For completeness, we include the comparison to the 
  automatic differentiation variational inference (ADVI) 
  algorithm in \texttt{rstan} \cite[]{kucukelbir_automatic_2015}. ADVI is a 
  variational Bayes algorithm that assumes a full factorization of the posterior
  and is fit through gradient based optimization of the variational objective.
  \texttt{rstan} offers an ADVI algorithm that assumes a fullrank posterior 
  covariance, but this algorithm failed often resulted in errors and
  convergence problems, so we do not include it in the comparisons.

  To assess the performance of the variational approximations we evaluate 
  several measures. Divergence between the MCMC and variational posterior
  is quantified through a multivariate extension of the Cram{\'e}r-von Mises
  criterion \cite[]{baringhaus_new_2004}, where smaller values of the criterion
  indicate more similar distributions. In addition, we calculate the correlation
  and mean squared distance (MSD) between the posterior variances of
  the MCMC posterior and variational posterior. \texttt{rstan} allows the user 
  to maximise the posterior to obtain the maximum \textit{a posteriori} (MAP)
  estimates of the model parameters. We include these MAP estimates together
  with the ridge estimates obtained from the \texttt{glmnet} \texttt{R} package
  \cite[]{friedman_regularization_2010} in the comparison of the posterior mean
  estimates.

  The interest of this simulation lies in the variational Bayes approximation, 
  so we restrict ourselves to the univariate setting (i.e., $D=1$) and drop any
  $d$ indices from notation in the following. We simulated
  $n=100$ features $\x_i$ from a $p=200$ dimensional Gaussian distribution with
  zero mean vector, variances equal to one, and all covariances set to 
  $\rho=0.5$. The outcomes are generated through $y_i = \x_i\tr \bbeta + 
  \epsilon_i$, with the $\epsilon_i$ from the standard normal distribution. 
  Likewise, we generate the model parameters $\beta_j$ from a standard normal
  distribution. For estimation, we set our hyperparameters $\lambda=0.1$ and 
  $\theta=2$. We choose the penalty parameter in \texttt{glmnet} by assuming a 
  known error variance and matching the marginal prior variance of the NIG and
  ridge models, which results in the penalty parameter 
  $1/(n\sigma^2\theta)=1/200$.To mitigate the effect of random variation, we
  repeat the simulation 100 times and report the results in Figures 
  \ref{fig:boxplots_igaussian_res3_vb}-
  \ref{fig:boxplots_igaussian_res3_best}.

<<boxplots_igaussian_res3_vb, fig.cap="(a) Cramér-von Mises criterion, (b) correlation, and (c) MSD between the variances of the MCMC and variational bayes posteriors.", out.width="80%", fig.asp=1/2>>=
@

<<boxplots_igaussian_res3_best, fig.cap="(a) Correlation and (b) MSD between the $\\beta_j$ point estimates of MCMC and ADVI, VB, glmnet, and MAP.", out.width="80%", fig.asp=3/5>>=
@

  \section{Empirical Bayes estimation}
  We simulate from a correctly specified model given in \ref{eq:priormodel}. We 
  assume random $y_{id}$, $\x_i$, and $\bm{\beta}_d$, while we fix $\mathbf{C}$,
  $\gamma_d^2$, $\sigma_d^2$ and $\bm{\alpha}$. We estimate the $\bm{\mu}_d$,
  $\bm{\Sigma}_d$, $\delta_d$, $\bm{\alpha}$ and $\lambda$, while we fix the
  $\sigma_d^2$ to their true values.

  We set $n=100$, $p=200$, and $D=10$. We consider 5 evenly sized classes of
  drugs, such that $\mathbf{c}_{kd} = \mathbbm{1}(k = \text{class}_d)$ and
  $\begin{bmatrix} \mathbf{c}_1 & \cdots & \mathbf{c}_D \end{bmatrix} \tr$. From
  these the $\gamma_d^2 = (\mathbf{c}_d \tr \bm{\alpha})^{-1}$ are created. 
  Next, we simulate $\bm{\beta}_{jd} \sim \mathcal{N} 
  (0, \sigma^2_d \gamma_d^2)$. We fix the mean signal-to noise ratio 
  $\overline{\text{SNR}}_d$, such that the predictor data variance is 
  $s^2 = \V(\mathbf{x}_i) = \overline{\text{SNR}}_d / 
  (\overline{\gamma^2_d} p)$ (NOT CORRECT). The last step consists of 
  simulating the data by $\mathbf{x}_{ij} \sim \mathcal{N}(0,s^2)$ and 
  $y_{id} = \mathcal{N} (\mathbf{x}_i \tr \bm{\beta}_d, \sigma_d^2)$. The
  remaining parameters are set as follows: $\bm{\alpha} = \begin{bmatrix} 1 &
  \dots & 5 \end{bmatrix} \tr$, $\sigma_d^2 = 1$ for all $d$, and 
  $\overline{\text{SNR}}_d = 10$. To account for the random data and parameter
  generation, we repeat the simulation 100 times. We compare the results to the
  inverse Gamma and non-conjugate model, and present the results in Figures
  \ref{fig:boxplot_igaussian_res1_prior_mean}-
  \ref{fig:boxplot_igaussian_res2_prior_mean}.

@@

  - investigate convergence: (i) look at both error and prior variance 
  convergence and (ii) look at actual prior distribution convergence (
  plot true distribution and estimated prior every 5 iterations or so)

  - try to fix convergence: (i) track only expectation and variance or ELBO, 
  (ii) reparametrise, and (iii) fix $\forall d: \lambda_d=\lambda$.

@@

  \section{Discussion}
  - inverse Gaussian more convenient than inverse Gamma, because EB is closed-
  form

  - use copula's and marginal posteriors to find multivariate posteriors for
  contrast testing.

  - apart from the scaling issue, the conjugate prior model is preferred
  over the non-conjugate model in \cite{moran_variance_2018}.
  The assumed factorization in the variational Bayes approximation 
  $q(\bbeta_d) \cdot q(\sigma_d)$ is actually true for this model. The 
  approximate expectation used in the empirical Bayes step is therefore more 
  accurate
